---
  title: FDR estimation (Bottomly et. al.)
  output:
    html_document:
      fig_height: 15
      fig_width: 10
---

# transcript level
## loading data

Take a look at `run_benchmarks.R` to see how this data is generated.

```{r}
suppressMessages({
library('ggplot2')
library('mamabear')
library('dplyr')
library('cowplot')
})
```

```{r}
theme_hp <- function() {
  theme_cowplot(25) +
    theme(legend.key.size = unit(2, "lines"), legend.position = 'none')
}

base_dir <- '../results/final_figures'
default_extension <- '.pdf'
source('../../geuvadis/R/gene_common.R')
```


```{r}
self_benchmark <- readRDS('../results/isoform_self_benchmark.rds')
```

## creating the plots

```{r}
# debugonce(mamabear:::get_sensitivity_specificity)
self_fdr <- lapply(self_benchmark, average_sensitivity_specificity)
self_fdr <- dplyr::bind_rows(self_fdr)
self_fdr <- dplyr::mutate(self_fdr, method = sub('qval_', '', method))
```

Let’s rename things to be a bit more consistent with the main figures:

```{r}
mapping <- c('sleuth.lrt' = 'sleuth', 'limmaVoom' = 'voom',
  'DESeq2' = 'DESeq2', 'DESeq' = 'DESeq', 'Cuffdiff2' = 'Cuffdiff 2',
  'EBSeq' = 'EBSeq', 'edgeR' = 'edgeR')
self_fdr <- dplyr::filter(self_fdr, method != 'sleuth.wt')
self_fdr <- dplyr::mutate(self_fdr, method = mapping[method])
```

Now let’s make some pretty labels:

```{r}
self_fdr <- dplyr::mutate(self_fdr,
  fdr_level_string = paste0('eFDR = ', sprintf('%.2f', fdr_level)))
```

count the number of observations at each eFDR

```{r}
fdr_count <- dplyr::select(self_fdr, method, fdr_level_string, true_fdr, sample) %>%
  dplyr::group_by(method, fdr_level_string) %>%
  dplyr::summarize(n = sum(!is.na(true_fdr)))
fdr_count <- dplyr::mutate(fdr_count, y = 1)
```

```{r}
p <- ggplot(self_fdr, aes(method, true_fdr, color = method)) +
  geom_boxplot(outlier.shape = NA) + #avoid plotting outliers twice
  geom_jitter(position = position_jitter(width = 0.7, height = 0), alpha = 0.5) +
  geom_text(aes(method, y, label = n), data = fdr_count) +
  facet_wrap(~fdr_level_string, ncol = 1) +
  geom_hline(aes(yintercept = fdr_level), linetype = 2) +
  ylim(0, 1) +
  ylab('false discovery rate') +
  theme_hp() +
  scale_color_manual(values = method_colors)
p
```

```{r}
filename <- file.path(base_dir, paste0('isoform_resampling', default_extension))
save_plot(filename, p, base_aspect_ratio = 0.75, base_height = 15)
```

Save the results for later:

```{r}
saveRDS(self_fdr, file.path('..', 'results', 'self_fdr_isoform.rds'))
```

# gene level

## loading data
Take a look at run_benchmarks_gene.R to see how this data is generated.

```{r}
self_benchmark <- readRDS('../results/gene_self_benchmark.rds')
```

## creating the plots

```{r}
self_fdr <- lapply(self_benchmark, average_sensitivity_specificity)
self_fdr <- dplyr::bind_rows(self_fdr)
self_fdr <- dplyr::mutate(self_fdr, method = sub('qval_', '', method))
```

Let’s rename things to be a bit more consistent with the main figures:

```{r}
mapping <- c('sleuth.lrt' = 'sleuth', 'limmaVoom' = 'voom',
  'DESeq2' = 'DESeq2', 'DESeq' = 'DESeq', 'Cuffdiff2' = 'Cuffdiff 2',
  'EBSeq' = 'EBSeq', 'edgeR' = 'edgeR')
self_fdr <- dplyr::filter(self_fdr, method != 'sleuth.wt')
self_fdr <- dplyr::mutate(self_fdr, method = mapping[method])
```

Now let’s make some pretty labels:

```{r}
self_fdr <- dplyr::mutate(self_fdr,
  fdr_level_string = paste0('eFDR = ', sprintf('%.2f', fdr_level)))
```

count the number of observations at each eFDR

```{r}
fdr_count <- dplyr::select(self_fdr, method, fdr_level_string, true_fdr, sample) %>%
  dplyr::group_by(method, fdr_level_string) %>%
  dplyr::summarize(n = sum(!is.na(true_fdr)))
fdr_count <- dplyr::mutate(fdr_count, y = 1)
```

```{r}
p <- ggplot(self_fdr, aes(method, true_fdr, color = method)) +
  geom_boxplot(outlier.shape = NA) + #avoid plotting outliers twice
  geom_jitter(position = position_jitter(width = 0.7, height = 0), alpha = 0.5) +
  geom_text(aes(method, y, label = n), data = fdr_count) +
  facet_wrap(~fdr_level_string, ncol = 1) +
  geom_hline(aes(yintercept = fdr_level), linetype = 2) +
  ylim(0, 1) +
  ylab('false discovery rate') +
  theme_hp() +
  scale_color_manual(values = method_colors)
p
```

```{r}
filename <- file.path(base_dir, paste0('gene_resampling', default_extension))
save_plot(filename, p, base_aspect_ratio = 0.75, base_height = 15)
```

let's save everything:

```{r}
saveRDS(self_fdr, file.path('..', 'results', 'self_fdr_gene.rds'))
```
