---
  title: "transcript level benchmarks (Bottomly et. al.)"
  output:
    html_document:
      fig_height: 14
---

# preliminaries

We must first load the data.
To see how the comparisons were generated, look at `run_benchmarks.R`.

```{r}
# all_benchmarks <- readRDS('../results/all_benchmarks.rds')
sensitivity_precision_plots <- readRDS('../results/sensitivity_precision_plots.rds')
```

## organization

`all_benchmarks` and `sensitivity_precision_plots` are organized in the following manner:

- first level: the "truth"
- second level: the method being compared
- third level: the replications of the second level

## loading some packages

```{r}
library('ggplot2')
library('mamabear')
library('cowplot')
```

# sensitivity and precision plots while rotating different methods

## sleuth likelihood ratio test as the validation set

```{r}
current_validation <- sensitivity_precision_plots$sleuth.lrt
```

### limmaVoom

```{r}
current_validation$
  limmaVoom$
  together
```

### DESeq2

```{r}
current_validation$
  DESeq2$
  together
```

### EBSeq

```{r}
current_validation$
  EBSeq$
  together
```

## sleuth wald test as the validation set

```{r}
current_validation <- sensitivity_precision_plots$sleuth.wt
```

### limmaVoom

```{r}
current_validation$
  limmaVoom$
  together
```

### DESeq2

```{r}
current_validation$
  DESeq2$
  together
```

### EBSeq

```{r}
current_validation$
  EBSeq$
  together
```

## limmaVoom as the validation set

```{r}
current_validation <- sensitivity_precision_plots$limmaVoom
```

### limmaVoom

```{r}
current_validation$
  limmaVoom$
  together
```

### DESeq2

```{r}
current_validation$
  DESeq2$
  together
```

### EBSeq

```{r}
current_validation$
  EBSeq$
  together
```

## DESeq2 as the validation set

```{r}
current_validation <- sensitivity_precision_plots$DESeq2
```

### limmaVoom

```{r}
current_validation$
  limmaVoom$
  together
```

### DESeq2

```{r}
current_validation$
  DESeq2$
  together
```

### EBSeq

```{r}
current_validation$
  EBSeq$
  together
```

## EBSeq as the validation set

```{r}
current_validation <- sensitivity_precision_plots$EBSeq
```

### limmaVoom

```{r}
current_validation$
  limmaVoom$
  together
```

### DESeq2

```{r}
current_validation$
  DESeq2$
  together
```

### EBSeq

```{r}
current_validation$
  EBSeq$
  together
```
