---
  title: "looking at different variance components"
  output:
    html_document:
      fig_height: 20
      fig_width: 20
---

# preliminaries

Load things like the transcript to gene mapping and permutation info.

```{r}
source('../../geuvadis/R/gene_common.R')
source('../../geuvadis/R/benchmark_methods.R')
source('get_metadata.R')
library('data.table')
library('cowplot')
```

```{r}
theme_hp <- function() {
  theme_cowplot(25) +
    theme(legend.key.size = unit(2, "lines"), legend.position = 'none')
}
default_extension <- '.pdf'
base_dir <- file.path('..', 'results', 'final_figures', 'technical_variation')
if (!dir.exists(base_dir)) {
  dir.create(base_dir)
}
```

```{r}
set.seed(42)
which_permutation <- sample.int(20, 1)
```

```{r}
stc <- training_sets[[which_permutation]]
```

# transcript level

```{r}
isoform_file <- '../results/sleuth_isoform.rds'

if (!file.exists(isoform_file)) {
  soi <- run_sleuth_prep(stc)
  soi <- sleuth_fit(soi, ~1, 'reduced')
  soi <- sleuth_lrt(soi, 'reduced', 'full')
  saveRDS(soi, file = isoform_file)
} else {
  soi <- readRDS(isoform_file)
}
```

get the relevant results out

```{r}
sr <- sleuth_results(soi, 'reduced:full', 'lrt', show_all = FALSE)
```

```{r}
isoform_alpha = 0.15
```

plot the total raw variance

```{r}
ggplot(sr, aes(mean_obs, (var_obs) ^ (0.25))) +
  geom_point(alpha = isoform_alpha) +
  theme_cowplot(25) +
  xlab('mean( log( counts + 0.5 ) )') +
  ylab('sqrt( total standard deviation )')
```

plot the biological variance with shrinkage:

```{r}
plot_mean_var(soi) +
  theme_cowplot(25)
```

plot the technical variability:

```{r}
ggplot(sr, aes(mean_obs, (tech_var) ^ 0.25)) +
  geom_point(alpha = 0.05) +
  theme_cowplot(25) +
  xlab('mean( log( counts + 0.5 ) )') +
  ylab('sqrt( inferential standard deviation )')
```

plot the final variability estimates (total)

```{r}
ggplot(sr, aes(mean_obs, (final_sigma_sq + tech_var) ^ 0.25)) +
  geom_point(alpha = 0.05) +
  theme_cowplot(25) +
  xlab('mean( log( x + 0.5 ) )') +
  ylab('sqrt( final standard deviation )')
```

# gene level analysis

```{r}
gene_file <- '../results/sleuth_gene.rds'

sog <- NULL
if (!file.exists(gene_file)) {
  sog <- run_sleuth_prep(stc, gene_mode = 'ens_gene')
  sog <- sleuth_fit(sog, ~1, 'reduced')
  sog <- sleuth_lrt(sog, 'reduced', 'full')
  saveRDS(sog, file = gene_file)
} else {
  sog <- readRDS(gene_file)
}
```

get the relevant results out

```{r}
sr <- sleuth_results(sog, 'reduced:full', 'lrt', show_all = FALSE)
```


some global constants

```{r}
gene_alpha <- 0.7
```

plot the total raw variance

```{r}
ggplot(sr, aes(mean_obs, (var_obs) ^ (0.25))) +
  geom_point(alpha = gene_alpha) +
  theme_cowplot(25) +
  xlab('mean( log( x + 0.5 ) )') +
  ylab('sqrt( total standard deviation )')
```

plot the biological variance with shrinkage:

```{r}
plot_mean_var(sog) +
  theme_cowplot(25)
```

plot the technical variability:

```{r}
ggplot(sr, aes(mean_obs, (tech_var) ^ 0.25)) +
  geom_point(alpha = gene_alpha) +
  theme_cowplot(25) +
  xlab('mean( log( x + 0.5 ) )') +
  ylab('sqrt( inferential standard deviation )')
```

plot the final variability estimates (total)

```{r}
ggplot(sr, aes(mean_obs, (final_sigma_sq + tech_var) ^ 0.25)) +
  geom_point(alpha = gene_alpha) +
  theme_cowplot(25) +
  xlab('mean( log( x + 0.5 ) )') +
  ylab('sqrt( final standard deviation )')
```


```{r}
estimate_transformed_variance <- function(mu, effective_length, n, scale_factor) {
  # compute something proportional to TPM
  res <- sapply(1:5000, function(i) {
    y <- rpois(n, mu) / effective_length
    y <- log(scale_factor * y)
    y <- y[is.finite(y)]
    c(mean(y, na.rm = TRUE), var(y, na.rm = TRUE))
  })
  t(res)
}

# mimic sampling the same number of things that you would in the kallisto bootstrap
n_bootstrap <- length(sog$kal[[1]]$bootstrap)

set.seed(42)
system.time(tmp <- lapply(exp(seq(log(0.5), log(10000000), length.out = 3000)),
  function(mu) {
    estimate_transformed_variance(mu, 1, n_bootstrap, 1)
  })
)
```

```{r}
confidence_bands <- lapply(tmp,
  function(x) {
    # compute the lower and upper confidence intervals of the variance estimate
    qs <- quantile(x[, 2], probs = c(0.05, 0.95, 0.999, 0.25, 0.5, 0.75),
      na.rm = TRUE)
    data.frame(mean_obs = mean(x[, 1], na.rm = TRUE),
      lower = qs[1], upper = qs[2], upper_extreme = qs[3],
      mean_var = mean(x[, 2], na.rm = TRUE),
      q25 = qs[4], q50 = qs[5], q75 = qs[6]
      )
  })
confidence_bands <- dplyr::bind_rows(confidence_bands)
confidence_bands <- dplyr::filter(confidence_bands, is.finite(mean_obs))
confidence_bands <- dplyr::mutate(confidence_bands, iqr = q75 - q25)
```

a sanity check on the confidence bands to ensure that we are sampling reasonably:

```{r}
ggplot(confidence_bands, aes(mean_obs, mean_var)) +
  geom_point() +
  geom_point(aes(mean_obs, lower), color = 'red') +
  geom_point(aes(mean_obs, upper), color = 'red') +
  geom_point(aes(mean_obs, upper_extreme))
```

## simply looking at 1 replicate

```{r}
bs_summary <- sleuth:::sleuth_summarize_bootstrap_col(sog, 'est_counts',
  function(x) log(x + 0.5))
```

```{r}
obj_mod <- sog

transform <- function(x) log(x + 0.5)

which_column <- 'ens_gene'
obj_mod$target_mapping <- dplyr::mutate(obj_mod$target_mapping,
  target_id = sapply(strsplit(target_id, '\\.'), '[[', 1))

obj_mod$obs_norm <- dplyr::mutate(obj_mod$obs_norm,
  target_id = sapply(strsplit(target_id, '\\.'), '[[', 1))
obj_mod$obs_norm_filt <- dplyr::mutate(obj_mod$obs_norm_filt,
  target_id = sapply(strsplit(target_id, '\\.'), '[[', 1))
obj_mod$obs_raw <- dplyr::mutate(obj_mod$obs_raw,
  target_id = sapply(strsplit(target_id, '\\.'), '[[', 1))

# scale_factor <- median(obj_mod$obs_norm_filt$eff_len)
# begin by gene median
tmp <- obj_mod$obs_raw
# tmp <- as.data.table(tmp)
tmp <- dplyr::left_join(
  data.table::as.data.table(tmp),
  data.table::as.data.table(obj_mod$target_mapping),
  by = 'target_id')

tmp <- dplyr::group_by_(tmp, 'sample', which_column)
scale_factor <- dplyr::mutate(tmp, scale_factor = median(eff_len))
# end by gene median
obj_mod$obs_norm_filt <- sleuth:::reads_per_base_transform(obj_mod$obs_norm_filt,
  scale_factor, which_column, obj_mod$target_mapping, norm_by_abundance = TRUE)
obj_mod$obs_norm <- sleuth:::reads_per_base_transform(obj_mod$obs_norm,
  scale_factor, which_column, obj_mod$target_mapping, norm_by_abundance = TRUE)

obs_counts <- sleuth:::obs_to_matrix(obj_mod, "scaled_reads_per_base")
obs_counts <- transform(obs_counts)

obj_mod$kal <- lapply(seq_along(obj_mod$kal),
  function(i) {
    k <- obj_mod$kal[[i]]
    current_sample <- obj_mod$sample_to_covariates$sample[i]
    print(current_sample)
    k$bootstrap <- lapply(k$bootstrap, function(b) {
      b <- dplyr::mutate(b,
        target_id = sapply(strsplit(target_id, '\\.'), '[[', 1))
      b <- dplyr::mutate(b, sample = current_sample)
      sleuth:::reads_per_base_transform(b, scale_factor, which_column,
        obj_mod$target_mapping, norm_by_abundance = TRUE)
    })

    k
  })

bs_summary <- sleuth:::sleuth_summarize_bootstrap_col(obj_mod, "scaled_reads_per_base",
  transform)
```

```{r}
which_sample <- 'SRR099228'
bs_summary <- dplyr::filter(bs_summary, sample == which_sample)
obs_counts_filter <- obs_counts[sr$target_id, ]
bs_summary <- dplyr::semi_join(bs_summary, sr, by = 'target_id')

bs_summary <- dplyr::select(bs_summary, target_id,
  bs_var = bs_var_scaled_reads_per_base)
em_counts <- data.frame(target_id = rownames(obs_counts_filter),
  est_counts = obs_counts_filter[, which_sample])
bs_summary <- dplyr::left_join(bs_summary, em_counts, by = 'target_id')

bs_summary <- dplyr::filter(bs_summary, est_counts > log(0.5))
```

```{r}
bs_summary <- dplyr::mutate(bs_summary,
  cv_estimate = bs_var / est_counts ^ 2,
  cv_rank = ecdf(cv_estimate)(cv_estimate),
  cv_top = cv_rank >= 0.90)
```

perform a rolling join trying to find the nearest simulation Poisson point to the mean of this value.
also, ensure that the relative error is less than 1%.

```{r}
bs_summary <- as.data.table(bs_summary)
setkey(bs_summary, est_counts)
confidence_bands <- as.data.table(confidence_bands)
setkey(confidence_bands, mean_obs)

bs_summary_tmp <- bs_summary
bs_summary_tmp <- bs_summary_tmp[, est_counts_new := est_counts]

bs_summary_roll <- confidence_bands[bs_summary_tmp, roll = 1,
  rollends = c(TRUE, TRUE), allow.cartesian = TRUE]

bs_summary_roll <- dplyr::rename(bs_summary_roll, est_counts = est_counts_new)

bs_summary_roll <- dplyr::mutate(bs_summary_roll,
  near_filter = ((est_counts - mean_obs) / est_counts) < 0.01)
```

now we can find outliers based on how far they are from the expected Poisson:

```{r}
bs_summary_roll <- dplyr::mutate(bs_summary_roll,
  outlier = (bs_var > upper_extreme) & near_filter)
```

use the "outlier rule" with varying multiplicative factors

```{r}
bs_summary_roll <- dplyr::mutate(bs_summary_roll,
  iqr_outlier_1.5 = (iqr * 1.5 + q75) < bs_var,
  iqr_outlier_2 = (iqr * 2 + q75) < bs_var,
  iqr_outlier_3 = (iqr * 3 + q75) < bs_var,
  iqr_outlier_10 = (iqr * 10 + q75) < bs_var,
  iqr_outlier_100 = (iqr * 100 + q75) < bs_var
  )
summary(bs_summary_roll)
```

some global constants

```{r}
gene_alpha <- 0.55
```

100 seems to give a reasonable number of outliers

plot inferential variability versus the mean counts

```{r}
p <- ggplot(bs_summary_roll, aes(est_counts, bs_var ^ 0.25)) +
  # geom_point(aes(color = outlier), alpha = gene_alpha) +
  geom_point(aes(color = iqr_outlier_100), alpha = gene_alpha) +
  scale_color_manual(values = c('black', 'red')) +
  xlab('log( abundance + 0.5 ) )') +
  ylab('sqrt( inferential standard deviation )') +
  geom_path(aes(mean_obs, lower ^ 0.25), color = 'springgreen',
    size = 0.8, alpha = 0.7, linetype = 1) +
  geom_path(aes(mean_obs, upper ^ 0.25), color = 'springgreen',
    size = 0.8, alpha = 0.7, linetype = 1) +
  geom_path(aes(mean_obs, mean_var ^ 0.25),
    color = 'springgreen', size = 1.2, alpha = 1) +
  theme_hp()
p
```

```{r}
filename <- file.path(base_dir, paste0('inferential_variance',
  default_extension))
save_plot(filename, p, base_aspect_ratio = 1.6, base_height = 15)
```


# run DESeq2 and get dispersion estimates

```{r}
training_sets <- lapply(training_sets,
  function(df) {
    df$featureCounts <- file.path('..', 'results', 'single', df$sample,
      'featureCounts.txt')
    df
  })

stc <- training_sets[[which_permutation]]
gene_names <- unique(transcript_gene_mapping$ens_gene)

dummy_filter <- rep(TRUE, length(gene_names))
names(dummy_filter) <- gene_names

training_counts <- lapply(training_sets,
  function(df) {
    obs <- load_union_counts_general(df$featureCounts, df$sample)
    current_filter <- intersect(rownames(obs), names(dummy_filter))
    obs[current_filter, ]
  })

current_counts <- training_counts[[which_permutation]]

cds <- make_count_data_set(round(current_counts), stc)
dds <- DESeqDataSetFromMatrix(exprs(cds), DataFrame(pData(cds)), ~ condition)
dds <- dds[ rowSums(counts(dds)) > 1,]
dds <- DESeq(dds)

plotDispEsts(dds)
```

```{r}
res <- as.data.frame(mcols(dds))
res <- dplyr::mutate(res, target_id = rownames(dds),
  final_dispersion = dispersions(dds))
res <- dplyr::filter(res, baseMean > 0)
```

join in information about coefficient of variation

```{r}
res <- dplyr::left_join(res,
  dplyr::select(bs_summary_roll, target_id, iqr_outlier_100),
  # dplyr::select(bs_summary_roll, target_id, iqr_outlier_100),
  by = 'target_id')
```

smooth variance

```{r}
moment_estimator <- ggplot(res,
  aes(log(baseMean), log(baseMean + dispGeneEst * baseMean ^ 2))) +
  geom_point(aes(color = iqr_outlier_100), alpha = gene_alpha) +
  # geom_point(aes(color = outlier), alpha = gene_alpha) +
  scale_color_manual(values = c('black', 'red')) +
  theme_cowplot(25) +
  xlab('log(mean( counts ))') +
  ylab('log(raw total variance)') +
  theme_hp()

smooth_variance <- ggplot(res,
  aes(log(baseMean),
    log(baseMean + final_dispersion * baseMean ^ 2))) +
  geom_point(aes(color = iqr_outlier_100), alpha = gene_alpha) +
  scale_color_manual(values = c('black', 'red')) +
  xlab('log(mean( counts ))') +
  ylab('log(final total variance)') +
  theme_hp()

p <- plot_grid(moment_estimator, smooth_variance)
p
filename <- file.path(base_dir, paste0('DESeq2_variance',
  default_extension))
save_plot(filename, p, base_aspect_ratio = 1.6, base_height = 15)
```

merge in the outlier information

```{r}
sr <- dplyr::left_join(sr,
  # dplyr::select(bs_summary_roll, target_id, outlier),
  dplyr::select(bs_summary_roll, target_id, iqr_outlier_100),
  by = 'target_id')
```

plot the total raw variance

```{r}
p <- ggplot(sr, aes(mean_obs, (var_obs) ^ (0.25))) +
  geom_point(aes(color = iqr_outlier_100), alpha = gene_alpha) +
  # geom_point(aes(color = outlier), alpha = gene_alpha) +
  scale_color_manual(values = c('black', 'red')) +
  theme_hp() +
  xlab('mean( log( abundance + 0.5 ) )') +
  ylab('sqrt( total standard deviation )')
p
filename <- file.path(base_dir, paste0('total_raw_variance',
  default_extension))
save_plot(filename, p, base_aspect_ratio = 1.6, base_height = 15)
```

plot the biological variance with shrinkage:

```{r}
tmp_summary <- dplyr::select(sog$fits[['full']]$summary, target_id, iqr)
sr <- dplyr::left_join(sr, tmp_summary, by = 'target_id')
```

```{r}
p <- ggplot(sr, aes(mean_obs, sqrt(sqrt(sigma_sq_pmax))))
p <- p + geom_point(aes(colour = iqr_outlier_100), alpha = gene_alpha,)
p <- p + geom_point(aes(color = iqr_outlier_100), alpha = gene_alpha)
p <- p + geom_line(aes(mean_obs, sqrt(sqrt(smooth_sigma_sq))),
    alpha = 1, size = 1, colour = 'deepskyblue')
p <- p + scale_colour_manual(values = c("black", "red"))
p <- p + theme(legend.position = "none")
p <- p + xlab("mean( log( abundance + 0.5 ) )")
p <- p + ylab("sqrt( biological variation )")
p <- p + theme_hp()
p
filename <- file.path(base_dir, paste0('biological_variance',
  default_extension))
save_plot(filename, p, base_aspect_ratio = 1.6, base_height = 15)
```


plot the final variability estimates (total)

```{r}
p <- ggplot(sr, aes(mean_obs, (final_sigma_sq + tech_var) ^ 0.25)) +
  geom_point(aes(color = iqr_outlier_100), alpha = gene_alpha) +
  # geom_point(aes(color = outlier), alpha = gene_alpha) +
  scale_color_manual(values = c('black', 'red')) +
  theme_hp() +
  xlab('mean( log( abundance + 0.5 ) )') +
  ylab('sqrt( final standard deviation )')
p
filename <- file.path(base_dir, paste0('final_biological_variance',
  default_extension))
save_plot(filename, p, base_aspect_ratio = 1.6, base_height = 15)
```
