---
title: "Estimating variance under measurement error model"
output:
  html_document:
    toc: true
    fig_width: 14
    fig_height: 10
---

# Load info

```{r}
library("cowplot")
library("sleuth")

base_dir <- "../results/3_3_1_1"
```

```{r,cache=TRUE}
# de_info contains the 'true' fold change as well as which transcripts are DE
de_info <- read.table(gzfile(file.path(base_dir, "de_info.tsv.gz")),
  header = TRUE, stringsAsFactors = FALSE)
```

```{r}
sobj_fname <- "../tmp/sobj.rds"
sobj <- NULL
{
  if ( !file.exists(sobj_fname) ) {
    kal_fnames <- file.path(base_dir, "sample_1", 1:6, "kallisto", "abundance.h5" )


    kal <- lapply(kal_fnames, read_kallisto_h5, read_bootstrap = TRUE)

    s2c <- data.frame(sample = paste0("sample_", 1:6),
      condition = rep(c("A", "B"), each = 3), stringsAsFactors = FALSE)

    sobj <- new_sleuth(kal, s2c, ~ condition)
    rm(kal)
    gc()

    # let's treat it as we know the truth
    sobj$obs_norm <- sobj$obs_raw
    saveRDS(sobj, sobj_fname)
  } else {
    sobj <- readRDS(sobj_fname)
  }
}
```

```{r,cache=TRUE}
pass_filt <- sobj$obs_norm %>%
  group_by(target_id) %>%
  summarise( count_filt = sum(est_counts > 5) > 0.8 * n())
```

Fit the pooled variance measurement error model

```{r}
mev <- NULL
{
  mev_fname <- "../tmp/mev.rds"
  if( !file.exists(mev_fname) ){
    mev <- me_equal_var(sobj, pass_filt, function(x) log(x + 0.5))
    saveRDS( mev, mev_fname )
  } else {
    mev <- readRDS( mev_fname )
  }
}
```

Competing methods:
```{r}
deseq_res <- readRDS("../tmp/DESeq2.rds")
limma_res <- readRDS("../tmp/limma.rds")
edgeR_res <- readRDS("../tmp/edgeR.rds")
```

```{r}
deg_free <- 4
X <- sobj$design_matrix
A <- solve( t(X) %*% X )
```

Get the betas into a data.frame

```{r,cache=TRUE}
b <- sapply(mev$mes,
  function(x) {
    x$ols_fit$coefficients[2]
  })
names(b) <- names(mev$mes)
b <- data.frame(target_id = names(b), b = b)
rownames(b) <- NULL
```

# Different variance estimators

Let's add a column that is the "standard" OLS estimate of sigma (assuming no
ME)

```{r}
mev$summary <- mutate(mev$summary, sigma_sq_ols = rss / deg_free)
```

## EDA

Plot of technical variance vs OLS variance (RSS/(n-p)). In the ME model, the
estimate of sigma is RSS/(n-p) - sigma\_q\_sq:

```{r,cache=TRUE}
ggplot(mev$summary, aes(log2(sigma_q_sq), log2(sigma_sq_ols))) +
  stat_binhex(bins = 150) +
  geom_abline(intercept = 0, slope = 1, colour = 'red')
```

How many things in the ME model will end up with sigma < 0?

```{r}
mev$summary %>%
  group_by( sigma_sq <= 0 ) %>%
  summarise(n = n())
```

How does the sleuth ME model change the variances (compared to OLS)?

```{r,cache=TRUE}
ggplot(mev$summary, aes(log2(smooth_sigma_sq_pmax + sigma_q_sq), log2(sigma_sq_ols))) +
  stat_binhex(bins = 150) +
  geom_abline(intercept = 0, slope = 1, colour = 'red')
```

What does the new smooth variance look like versus the technical variance?

```{r,cache=TRUE}
ggplot(mev$summary, aes(log2(smooth_sigma_sq_pmax + sigma_q_sq), log2(sigma_q_sq))) +
  stat_binhex(bins = 150) +
  geom_abline(intercept = 0, slope = 1, colour = 'red')
```

## Meta

Function to compute t-statistic / p-values given a DF with the columns

```{r}
# assumes the A matrix is defined somewhere
compute_test <- function(df, betas, bio_var_str, tech_var_str = NULL) {
  stopifnot( nrow(df) == nrow(betas) )

  betas <- mutate(betas, target_id = as.character(target_id))
  df <- mutate(df, target_id = as.character(target_id))

  df <- df[match(betas$target_id, df$target_id), ]
  stopifnot( all.equal(df$target_id, betas$target_id) )

  sigma <- df[[bio_var_str]]
  if (!is.null(tech_var_str)) {
    cat("Adding in technical var\n")
    sigma <- sigma + df[[tech_var_str]]
  }
  se <- sqrt(sigma * A[2,2])
  betas <- mutate(betas,
    se_b = se,
    t_stat = b / se_b,
    pval = 2 * pt(abs(t_stat), deg_free, lower.tail = FALSE),
    pval_wald = 2 * pnorm(abs(t_stat), lower.tail = FALSE)
    )
  betas <- betas[complete.cases(betas),]
  betas$qval <- p.adjust(betas$pval, method = "BH")

  betas
}
```

## OLS

```{r}
ols_res <- compute_test(mev$summary, b, "sigma_sq_ols")
```

```{r,cache=TRUE}
ggplot(ols_res, aes(pval)) +
  geom_histogram(binwidth = 0.01) +
  geom_vline(xintercept = 0.10)
```

```{r,cache=TRUE}
ggplot(ols_res, aes(qval)) +
  geom_histogram(binwidth = 0.01) +
  geom_vline(xintercept = 0.10)
```

## ME shrinkage

```{r}
me_shrink <- compute_test(mev$summary, b, "smooth_sigma_sq_pmax", "sigma_q_sq")
me_shrink_wald <- mutate(me_shrink, pval = pval_wald, qval = p.adjust(pval,
    method = "BH"))
```

```{r,cache=TRUE}
ggplot(me_shrink, aes(pval)) +
  geom_histogram(binwidth = 0.01) +
  geom_vline(xintercept = 0.10)
```

```{r,cache=TRUE}
ggplot(me_shrink, aes(qval)) +
  geom_histogram(binwidth = 0.01) +
  geom_vline(xintercept = 0.10)
```

```{r,cache=TRUE}
ggplot(me_shrink_wald, aes(pval)) +
  geom_histogram(binwidth = 0.01) +
  geom_vline(xintercept = 0.10)
```

```{r,cache=TRUE}
ggplot(me_shrink_wald, aes(qval)) +
  geom_histogram(binwidth = 0.01) +
  geom_vline(xintercept = 0.10)
```

## ME shrinkage when sigma > 0

For this test, simply shrink when sigma > 0.

```{r}
mev$summary <- mutate(mev$summary,
  smooth_sigma_sq_gt0 = ifelse(sigma_sq <= 0, 0, smooth_sigma_sq_pmax)
  )
```

```{r}
me_shrink_gt0 <- compute_test(mev$summary, b, "smooth_sigma_sq_gt0", "sigma_q_sq")
```

## Don't add back in technical var

```{r}
me_no_tech <- compute_test(mev$summary, b, "smooth_sigma_sq_pmax")
```

## Scale variance

```{r}
me_shrink_scale <- me_shrink
me_shrink_scale <- mutate(me_shrink_scale, se_b = se_b / 2)
me_shrink_scale <- mutate(me_shrink_scale,
  t_stat = b / se_b,
  pval = 2 * pt(abs(t_stat), deg_free, lower.tail = FALSE))
me_shrink_scale$qval <- p.adjust(me_shrink_scale$pval, method = "BH")
```

# pFDR

```{r}
# gamma in the Storey paper
gs <- seq(0.02, 1, length.out = 30)
pi0 <- sapply(gs, function(g) with(me_shrink, sum(pval > g) / (length(pval) * (1 - g))))
pi0
plot(gs, pi0)
```

The true pi0:

```{r}
pi0_true <- NULL
{
  me_shrink_true<- semi_join(de_info, me_shrink, by = 'target_id')
  pi0_true <- with(me_shrink_true, sum( !is_de ) / length(is_de))
}
pi0_true
```

## implementation

```{r}
pfdr <- function(pvals, lam) {
  m <- length(pvals)
  pi0 <- sum( pvals > lam ) / ((1 - lam) * m)
  print(pi0)

  # let's just give it the true prop
  #pi0 <- pi0_true
  print(pi0)
  ordr <- order(pvals)
  p <- pvals[ordr]
  q <- p / 1:m
  q <- q * pi0 * m
  q_next <- c(q[2:m], q[m])
  q <- pmin(q, q_next)

  q[order(ordr)]
}

```

```{r}
me_shrink_pfdr <- mutate(me_shrink, qval = pfdr(pval, 0.5))
me_shrink_wald_pfdr <- mutate(me_shrink_wald, qval = pfdr(pval, 0.6))
```

```{r}
# gamma in the Storey paper
gs <- seq(0.02, 1, length.out = 30)
pi0 <- sapply(gs, function(g) with(me_shrink_wald, sum(pval > g) / (length(pval) * (1 - g))))
pi0
plot(gs, pi0)
```

```{r}
ggplot(me_shrink_pfdr, aes(qval)) +
  geom_histogram(binwidth = 0.01) +
  geom_vline(xintercept = 0.10) +
  xlim(0, 1)
```

```{r}
me_shrink_pfdr %>%
  filter(qval <= 0.10) %>%
  summarise(sig = n())
```

```{r}
ggplot(me_shrink_wald_pfdr, aes(qval)) +
  geom_histogram(binwidth = 0.01) +
  geom_vline(xintercept = 0.10) +
  xlim(0, 1)
```

```{r}
me_shrink_wald_pfdr %>%
  filter(qval <= 0.10) %>%
  summarise(sig = n())
```

# Benchmarks

```{r}
de_bench <- new_de_benchmark(
  list(
    deseq_res,
    limma_res,
    edgeR_res,
    ols_res,
    me_shrink,
    #me_shrink_gt0,
    #me_no_tech,
    #me_shrink_scale,
    #me_shrink_pfdr,
    me_shrink_wald,
    me_shrink_wald_pfdr
    ),
  c(
    "DESeq2",
    "voom",
    "edgeR (tagwise)",
    "OLS",
    "sleuth (equal var)",
    #"sleuth (equal var gt0)",
    #"sleuth (no tech)",
    #"sleuth (scale shrink)",
    #"sleuth (equal var, pFDR)",
    "sleuth (equal var, Wald)",
    "sleuth (equal var, Wald, pFDR)"
    ), de_info)
```

```{r}
fdr_nde_plot(de_bench, TRUE) +
  xlim(2000, 7500) +
  ylim(0, 0.2) +
  theme(legend.position = c(0.1, 0.85))
```

# Low technical variance

```{r}
mev_filt <- filter(mev$summary, sigma_sq > 0)
b_filt <- semi_join(b, mev_filt, by = 'target_id')
```

## ME shrinkage

```{r}
me_shrink_filt <- semi_join(me_shrink, mev_filt, by = 'target_id')
```

