---
  title: "benchmark report for simulations"
  output:
    html_document:
      fig_height: 20
      fig_width: 20
---

# preliminaries

```{r, eval=FALSE}
n_cpu <- 20
sim_name <- 'gfr_3_3_20_42_2'
sim_name <- 'gcd_3_3_20_1_2'
sim_name <- 'isoform_3_3_20_1_1'
```


```{r}
args <- commandArgs(trailingOnly = TRUE)
if (length(args) != 2) {
  stop('Usage: RScript N_CPU EXPERIMENT_STRING')
}
n_cpu <- args[1]
sim_name <- args[2]
```

The simulation string is:

```{r}
sim_name
```

We use the variable `simulation_mode` to build different axes depending on the simulation.
These axes have been tuned by hand.

```{r}
prefix <- strsplit(sim_name, '_')[[1]][1]
simulation_mode <- switch(prefix,
  'isoform' = 'independent',
  'gcd' = 'common',
  'gfr' = 'reference',
  stop(paste0('unrecognized simulation type. Please check the `sim_name`: ',
    sim_name)
  ))
```

The simulation mode is: `R simulation_mode`.

Some more global variables on where to save the figures:

```{r}
base_dir <- '../results/final_figures'
default_extension <- '.pdf'
```

A theme that seems to save well:

```{r}
theme_hp <- function() {
  theme_cowplot(25) +
    theme(legend.key.size = unit(2, "lines"))
}
```

## load packages

```{r}
suppressPackageStartupMessages({
library('cowplot')
library('data.table')
library('dplyr')
library('mamabear')
library('parallel')
})
source('gene_common.R')
source('benchmark_methods.R')
transcript_gene_mapping <- get_human_gene_names()
options(mc.cores = n_cpu)
```

set the number of columns for each grid

```{r}
n_grid_columns <- 2
```

ignore the following when plotting estimated fdr:

# gene benchmarks

## fdr/power curves when everyone is using their own filter

```{r}
each_filter_benchmark <- readRDS(paste0('../results/', sim_name,
  '/gene_benchmarks_filter.rds'))
```

```{r}
txi_each_filter <- readRDS(paste0('../results/', sim_name,
  '/txi.rds'))
```

Let's make a new benchmark object with just the ones we want:

```{r}
to_keep <- c('sleuth', 'edgeR', 'DESeq2', 'limmaVoom')

each_filter_benchmark <- lapply(seq_along(each_filter_benchmark),
  function(i) {
    sim_info <- get_de_info(sim_name, i, transcript_gene_mapping)

    original_data <- each_filter_benchmark[[i]]$original_data
    original_data <- original_data[to_keep]
    original_data$voom <- original_data$limmaVoom
    original_data$limmaVoom <- NULL
    new_data <- c(original_data,
      list(`DESeq2 (tximport)` = txi_each_filter$DESeq2[[i]]),
      list(`edgeR (tximport)` = txi_each_filter$edgeR[[i]]),
      list(`voom (tximport)` = txi_each_filter$voom[[i]]))

    new_de_benchmark(new_data, names(new_data), sim_info$de_genes,
      join_mode = 'union')
  })
```

because some of the true things will inevitably get filtered, remove the ones we think will be removed from the truth (`sim_filter = TRUE`).

```{r}
suppressMessages(current_fdr <- get_fdr(each_filter_benchmark)$pvals)
```

```{r}
method_colors <- method_colors[c('sleuth', 'DESeq2', 'edgeR', 'voom')]
method_colors <- c(method_colors, `DESeq2 (tximport)` = 'gray',
  `edgeR (tximport)` = '#c8acbc',
  `voom (tximport)` = '#e8cc90')
```


Zoom into a relevant area:

```{r}
tmp <- fdr_efdr_power_plot(current_fdr, start = 100, jump = 100, rank_fdr = 0.10,
  method_colors = method_colors, fdr_level_position = -0.005)

current_limits <- switch(simulation_mode,
  independent = list(x = c(-0.01, 0.25), y = c(-0.01, 0.25)),
  common = list(x = c(-0.01, 0.25), y = c(-0.01, 0.70)),
  reference = list(x = c(-0.01, 0.25), y = c(-0.01, 0.25))
  )

p <- tmp + theme_hp()
p <- p + coord_cartesian(xlim = current_limits$x, ylim = current_limits$y,
  expand = FALSE)
p <- p + guides(colour = guide_legend(override.aes = list(size=5)))
p

filename <- file.path(base_dir, paste0('txi_gene.each_filter_', sim_name,
  default_extension))
save_plot(filename, p, base_aspect_ratio = 1.6, base_height = 15)
```

```{r}
tmp <- fdr_efdr_power_plot(current_fdr, start = 500, jump = 500, rank_fdr = 0.10,
  method_colors = method_colors, fdr_level_position = -0.02)

p <- tmp +
  coord_cartesian(xlim = c(-0.05, 1), ylim = c(-0.05, 1), expand = FALSE) +
  theme_hp()
p <- p +
  geom_polygon(aes(x, y), alpha = 0.20,
    data = data.frame(
    x = c(0, 0, current_limits$x[2], current_limits$x[2]),
    y = c(0, current_limits$y[2], current_limits$y[2], 0)))
p <- p + guides(colour = guide_legend(override.aes = list(size=5)))
p

filename <- file.path(base_dir, paste0('txi_gene.each_filter_nozoom_', sim_name,
  default_extension))
save_plot(filename, p, base_aspect_ratio = 1.6, base_height = 15)
```

# Session Info

```{r}
sessionInfo()
```
