---
  title: "benchmark report for simulations"
  output:
    html_document:
      fig_height: 20
      fig_width: 20
---

# preliminaries

```{r, eval=FALSE}
sim_name <- 'gfr_3_3_20_42_2'
sim_name <- 'gcd_3_3_20_1_2'
sim_name <- 'isoform_3_3_20_1_1'
```


```{r}
args <- commandArgs(trailingOnly = TRUE)
if (length(args) != 2) {
  stop('Usage: RScript N_CPU EXPERIMENT_STRING')
}
n_cpu <- args[1]
sim_name <- args[2]
```

The simulation string is:

```{r}
sim_name
```

We use the variable `simulation_mode` to build different axes depending on the simulation.
These axes have been tuned by hand.

```{r}
prefix <- strsplit(sim_name, '_')[[1]][1]
simulation_mode <- switch(prefix,
  'isoform' = 'independent',
  'gcd' = 'common',
  'gfr' = 'reference',
  stop(paste0('unrecognized simulation type. Please check the `sim_name`: ',
    sim_name)
  ))
```

The simulation mode is: `R simulation_mode`.

Some more global variables on where to save the figures:

```{r}
base_dir <- '../results/final_figures'
default_extension <- '.pdf'
```

A theme that seems to save well:

```{r}
theme_hp <- function() {
  theme_cowplot(25) +
    theme(legend.key.size = unit(2, "lines"))
}
```

## load packages

```{r}
suppressPackageStartupMessages({
library('cowplot')
library('data.table')
library('dplyr')
library('mamabear')
library('parallel')
})
source('gene_common.R')
# options(mc.cores = n_cpu)
```

set the number of columns for each grid

```{r}
n_grid_columns <- 2
```

ignore the following when plotting estimated fdr:

```{r}
ignore_fdr <- c('GLFC', 'LFC')
method_colors <- c(method_colors, c(LFC = '#4393c3', GLFC = '#fddbc7',
  `sleuth (0 tech var)` = 'gray',
  `sleuth (Poisson tech var)` = 'coral',
  `sleuth (smooth technical var)` = 'olivedrab'
  ))
```

# isoform benchmarks

## fdr/power curves when everyone is using their own filter

```{r}
each_filter_benchmark <- readRDS(paste0('../results/', sim_name,
  '/isoform_benchmarks_filter_smooth_variance.rds'))
  # '/isoform_benchmarks_filter_zero_variance.rds'))
```

next, let's rename some things for aesthetics

```{r}
suppressMessages(
each_filter_benchmark <- lapply(each_filter_benchmark,
  rename_benchmark,
  c('Cuffdiff2', 'limmaVoom', 'sleuth_zero_variance', 'sleuth_poisson_variance',
    'sleuth_smooth_variance'),
  c('Cuffdiff 2', 'voom', 'sleuth (0 tech var)', 'sleuth (Poisson tech var)',
    'sleuth (smooth technical var)'), join_mode = 'union')
)
```

fixing weirdness that doesn't seem to matter because everyone filters out anyway

```{r}
if (simulation_mode == 'reference') {
  eps <- 0.005
  original_oracle <- lapply(each_filter_benchmark, '[[', 'oracle')
  original_data <- lapply(each_filter_benchmark,
    function(benchmark) {
      benchmark$original_data
    })
  efb <- each_filter_benchmark
  each_filter_benchmark <- lapply(seq_along(original_data),
    function(i) {
      cat(paste0('before: ', sum(original_oracle[[i]]$is_de), '\n'))
      current_oracle <- dplyr::mutate(original_oracle[[i]],
        is_de = ifelse(is_de & log_fc < eps & -eps < log_fc, FALSE, is_de))
      cat(paste0('after: ', sum(current_oracle$is_de), '\n'))

      new_de_benchmark(original_data[[i]], names(original_data[[i]]),
        current_oracle, de_colors = each_filter_benchmark[[i]]$color_mapping, join_mode = 'union')
    })
}
```

because some of the true things will inevitably get filtered, remove the ones we think will be removed from the truth (`sim_filter = TRUE`).

```{r}
suppressMessages(current_fdr <- get_fdr(each_filter_benchmark, sim_filter = TRUE)$pvals)
```

Zoom into a relevant area:

```{r}
tmp <- fdr_efdr_power_plot(current_fdr, start = 100, jump = 100, rank_fdr = 0.10,
  method_colors = method_colors, fdr_level_position = -0.005,
  ignore_estimated_fdr = ignore_fdr)

current_limits <- switch(simulation_mode,
  independent = list(x = c(-0.01, 0.25), y = c(-0.01, 0.28)),
  common = list(x = c(-0.01, 0.25), y = c(-0.01, 0.20)),
  reference = list(x = c(-0.01, 0.25), y = c(-0.01, 0.075))
  )

p <- tmp + theme_hp()
p <- p + coord_cartesian(xlim = current_limits$x, ylim = current_limits$y,
  expand = FALSE)
p

filename <- file.path(base_dir, paste0('0_variance_isoform.each_filter_', sim_name,
  default_extension))
save_plot(filename, p, base_aspect_ratio = 1.6, base_height = 15)
```

```{r}
tmp <- fdr_efdr_power_plot(current_fdr, start = 500, jump = 500, rank_fdr = 0.10,
  method_colors = method_colors, fdr_level_position = -0.02,
  ignore_estimated_fdr = ignore_fdr)

p <- tmp +
  coord_cartesian(xlim = c(-0.05, 1), ylim = c(-0.05, 1), expand = FALSE) +
  theme_hp()
p <- p +
  geom_polygon(aes(x, y), alpha = 0.20,
    data = data.frame(
    x = c(0, 0, current_limits$x[2], current_limits$x[2]),
    y = c(0, current_limits$y[2], current_limits$y[2], 0)))
p

filename <- file.path(base_dir, paste0('0_variance_isoform.each_filter_nozoom_', sim_name,
  default_extension))
save_plot(filename, p, base_aspect_ratio = 1.6, base_height = 15)
```

# Session Info

```{r}
sessionInfo()
```
