---
  title: "FDR estimation (resampling simulation et. al.)"
  output:
    html_document:
      fig_height: 12
      fig_width: 10
---

# preliminaries

```{r}
source('gene_common.R')
source('benchmark_methods.R')
```

## loading packages

```{r}
library('ggplot2')
library('mamabear')
library('dplyr')
library('cowplot')
```

```{r}
sim_name <- 'gfr_10_11_1_43_2'

theme_hp <- function() {
  theme_cowplot(25) +
    theme(legend.key.size = unit(2, "lines"), legend.position = 'none')
}
```

```{r}
base_dir <- '../results/final_figures'
default_extension <- '.pdf'
```

# getting the truth

Since this is a simulation, we can actually get the truth.

```{r}
transcript_gene_mapping <- get_human_gene_names()
sim_info <- get_de_info(sim_name, 1, transcript_gene_mapping)
```


# transcript level

## loading data

Take a look at `run_benchmarks.R` to see how this data is generated.

```{r}
self_benchmark <- readRDS(file.path('..', 'results', sim_name,
  'isoform_self_benchmark.rds'))
```

```{r}
self_benchmark_truth <- lapply(self_benchmark,
  function(method_list) {
    lapply(method_list, function (bench) {
      new_de_benchmark(bench$original_data, bench$labels, sim_info$de_info)
    })
  })
```

```{r}
tmp <- lapply(self_benchmark_truth,
  average_sensitivity_specificity, use_oracle = TRUE)
self_fdr_truth_df <- lapply(tmp,
  function(x) {
    x %>%
      dplyr::group_by(method, fdr_level) %>%
      dplyr::summarize(
        mean_fdr = mean(true_fdr, na.rm = TRUE),
        sd_fdr = sd(true_fdr, na.rm = TRUE))
  })
self_fdr_truth_df <- dplyr::bind_rows(self_fdr_truth_df)
```

# creating the plots


```{r}
self_fdr <- lapply(self_benchmark, average_sensitivity_specificity)
self_fdr <- dplyr::bind_rows(self_fdr)
self_fdr <- dplyr::mutate(self_fdr, method = sub('qval_', '', method))
```

Let's rename things to be a bit more consistent with the main figures:

```{r}
mapping <- c('sleuth.lrt' = 'sleuth', 'limmaVoom' = 'voom',
  'DESeq2' = 'DESeq2', 'DESeq' = 'DESeq', 'Cuffdiff2' = 'Cuffdiff 2',
  'EBSeq' = 'EBSeq', 'edgeR' = 'edgeR')
self_fdr <- dplyr::filter(self_fdr, method != 'sleuth.wt')
self_fdr <- dplyr::mutate(self_fdr, method = mapping[method])
self_fdr_truth_df <- dplyr::filter(self_fdr_truth_df, method != 'sleuth.wt')
self_fdr_truth_df <- dplyr::mutate(self_fdr_truth_df, method = mapping[method])
```

Now let's make some pretty labels:

```{r}
self_fdr <- dplyr::mutate(self_fdr,
  fdr_level_string = paste0('eFDR = ', sprintf('%.2f', fdr_level)))
self_fdr_truth_df <- dplyr::mutate(self_fdr_truth_df,
  fdr_level_string = paste0('eFDR = ', sprintf('%.2f', fdr_level)))
```

Let's put a position on top of each box plot:

```{r}
max_fdr_df <- dplyr::summarize(dplyr::group_by(self_fdr, method, fdr_level),
  max_fdr = max(true_fdr, na.rm = TRUE) + 0.04)
self_fdr_truth_df <- dplyr::inner_join(self_fdr_truth_df, max_fdr_df,
  by = c('fdr_level', 'method'))
self_fdr_truth_df <- dplyr::mutate(self_fdr_truth_df,
  label = sprintf('%.3f%s', mean_fdr,
    ifelse(!is.na(sd_fdr), sprintf(' \u00b1 %.3f', 2 * sd_fdr), '')))
```

Although, turns out were not going to do that anymore.
Here is the figure:

```{r}

p <- ggplot(self_fdr, aes(method, true_fdr, color = method)) +
  geom_boxplot() +
  # geom_text(aes(method, max_fdr + 0.02,
  #   label = label),
  #   data = self_fdr_truth_df, size = 6) +
  facet_wrap(~fdr_level_string, ncol = 1) +
  geom_hline(aes(yintercept = fdr_level), linetype = 2) +
  # ylim(0, 1) +
  coord_cartesian(ylim = c(0, 1.05)) +
  ylab('false discovery rate') +
  theme_hp() +
  geom_jitter() +
  scale_color_manual(values = method_colors)
p

filename <- file.path(base_dir, paste0('isoform_resampling', default_extension))
save_plot(filename, p, base_aspect_ratio = 0.75, base_height = 15)
```

save the transcript level stuff for a figure later:

```{r}
saveRDS(self_fdr, file = paste0('../results/', sim_name, '/self_fdr_isoform.rds'))
```

Note that voom behaves oddly at fdr = 0.01. This can be verified by looking at how many things it actually calls:

```{r}
lapply(seq_along(self_benchmark$limmaVoom),
  function(i) {
    x <- self_benchmark$limmaVoom[[i]]
    dplyr::filter(x$m_qval, estimate <= 0.01) %>%
      dplyr::mutate(sample = i)
  }) %>%
  dplyr::bind_rows()
```

### comparing self consistency and the true fdr

Since we have the actual fdr, let's compare the self consistency results with that.

```{r}
self_fdr_summary <- dplyr::group_by(self_fdr, method, fdr_level)
self_fdr_summary <- dplyr::summarize(self_fdr_summary,
  mean_fdr = mean(true_fdr, na.rm = TRUE),
  sd_fdr = sd(true_fdr, na.rm = TRUE))
self_fdr_summary <- dplyr::select(self_fdr_summary,
  method, fdr_level, self_mean_fdr = mean_fdr, self_sd_fdr = sd_fdr)
```

```{r}
fdr_summary <- dplyr::inner_join(self_fdr_truth_df, self_fdr_summary,
  by = c('method', 'fdr_level'))
fdr_summary <- dplyr::mutate(fdr_summary,
  fdr_low = mean_fdr - 2 * sd_fdr,
  fdr_high = mean_fdr + 2 * sd_fdr,
  self_fdr_low = self_mean_fdr - 2 * self_sd_fdr,
  self_fdr_high = self_mean_fdr + 2 * self_sd_fdr)
fdr_summary <- dplyr::mutate(fdr_summary,
  FDR = factor(sprintf('%.2f', fdr_level)))
```

```{r}
ggplot(fdr_summary, aes(self_mean_fdr, mean_fdr)) +
  geom_abline(intercept = 0, slope = 1, linetype = 3) +
  geom_point(aes(color = method, shape = FDR),
    size = 9, alpha = 0.8) +
  coord_cartesian(xlim = c(-0.05, 1), ylim = c(-0.05, 1), expand = FALSE) +
  scale_color_manual(values = method_colors) +
  geom_point(aes(x, y, shape = FDR), size = 5, color = 'black',
    data = data.frame(x = c(0.01, 0.05, 0.10), y = c(0.01, 0.05, 0.10),
      FDR = factor(c('0.01', '0.05', '0.10'))), show.legend = FALSE) +
  geom_errorbar(aes(ymax = fdr_high, ymin = fdr_low, color = method),
    width = 0.01) +
  # geom_errorbar(aes(xmax = self_fdr_high, xmin = self_fdr_low, color = method),
  #   width = 0.01) +
  # theme_hp() +
  theme_cowplot(25) +
  ylab('false discovery rate') +
  xlab('self-referential false discovery rate')
  # geom_vline()
```

# gene level

## loading data

Take a look at `run_benchmarks_gene.R` to see how this data is generated.

```{r}
self_benchmark <- readRDS(file.path('..', 'results', sim_name,
  'gene_self_benchmark.rds'))
```

Since this is a simulation, we know the truth.
Let's calculate true fdr:

```{r}
self_benchmark_truth <- lapply(self_benchmark,
  function(method_list) {
    lapply(method_list, function (bench) {
      new_de_benchmark(bench$original_data, bench$labels, sim_info$de_genes)
    })
  })
```

```{r}
tmp <- lapply(self_benchmark_truth,
  average_sensitivity_specificity, use_oracle = TRUE)
self_fdr_truth_df <- lapply(tmp,
  function(x) {
    x %>%
      dplyr::group_by(method, fdr_level) %>%
      dplyr::summarize(
        mean_fdr = mean(true_fdr, na.rm = TRUE),
        sd_fdr = sd(true_fdr, na.rm = TRUE))
  })
self_fdr_truth_df <- dplyr::bind_rows(self_fdr_truth_df)
```


# creating the plots

```{r}
self_fdr <- lapply(self_benchmark, average_sensitivity_specificity)
self_fdr <- dplyr::bind_rows(self_fdr)
self_fdr <- dplyr::mutate(self_fdr, method = sub('qval_', '', method))
```

Let's rename things to be a bit more consistent with the main figures:

```{r}
mapping <- c('sleuth.lrt' = 'sleuth', 'limmaVoom' = 'voom',
  'DESeq2' = 'DESeq2', 'DESeq' = 'DESeq', 'Cuffdiff2' = 'Cuffdiff 2',
  'EBSeq' = 'EBSeq', 'edgeR' = 'edgeR')
self_fdr <- dplyr::filter(self_fdr, method != 'sleuth.wt')
self_fdr <- dplyr::mutate(self_fdr, method = mapping[method])
self_fdr_truth_df <- dplyr::filter(self_fdr_truth_df, method != 'sleuth.wt')
self_fdr_truth_df <- dplyr::mutate(self_fdr_truth_df, method = mapping[method])
```

Now let's make some pretty labels:

```{r}
self_fdr <- dplyr::mutate(self_fdr,
  fdr_level_string = paste0('eFDR = ', sprintf('%.2f', fdr_level)))
self_fdr_truth_df <- dplyr::mutate(self_fdr_truth_df,
  fdr_level_string = paste0('eFDR = ', sprintf('%.2f', fdr_level)))
```

```{r}
max_fdr_df <- dplyr::summarize(dplyr::group_by(self_fdr, method, fdr_level),
  max_fdr = max(true_fdr, na.rm = TRUE) + 0.04)
self_fdr_truth_df <- dplyr::inner_join(self_fdr_truth_df, max_fdr_df,
  by = c('fdr_level', 'method'))
self_fdr_truth_df <- dplyr::mutate(self_fdr_truth_df,
  label = sprintf('%.3f%s', mean_fdr,
    ifelse(!is.na(sd_fdr), sprintf(' \u00b1 %.3f', 2 * sd_fdr), '')))
```

```{r}
p <- ggplot(self_fdr, aes(method, true_fdr, color = method)) +
  geom_boxplot() +
  # geom_text(aes(method, max_fdr + 0.02,
  #   label = label),
  #   data = self_fdr_truth_df, size = 6) +
  facet_wrap(~fdr_level_string, ncol = 1) +
  geom_hline(aes(yintercept = fdr_level), linetype = 2) +
  # ylim(0, 1) +
  coord_cartesian(ylim = c(0, 1.05)) +
  ylab('false discovery rate') +
  theme_hp() +
  geom_jitter() +
  scale_color_manual(values = method_colors)
p

filename <- file.path(base_dir, paste0('gene_resampling', default_extension))
save_plot(filename, p, base_aspect_ratio = 0.75, base_height = 15)
```

now let's save everything:

```{r}
saveRDS(self_fdr, file = paste0('../results/', sim_name, '/self_fdr_gene.rds'))
```

### comparing self consistency and the true fdr

Since we have the actual fdr, let's compare the self consistency results with that.

```{r}
self_fdr_summary <- dplyr::group_by(self_fdr, method, fdr_level)
self_fdr_summary <- dplyr::summarize(self_fdr_summary,
  mean_fdr = mean(true_fdr, na.rm = TRUE),
  sd_fdr = sd(true_fdr, na.rm = TRUE))
self_fdr_summary <- dplyr::select(self_fdr_summary,
  method, fdr_level, self_mean_fdr = mean_fdr, self_sd_fdr = sd_fdr)
```

```{r}
fdr_summary <- dplyr::inner_join(self_fdr_truth_df, self_fdr_summary,
  by = c('method', 'fdr_level'))
fdr_summary <- dplyr::mutate(fdr_summary,
  fdr_low = mean_fdr - 2 * sd_fdr,
  fdr_high = mean_fdr + 2 * sd_fdr,
  self_fdr_low = self_mean_fdr - 2 * self_sd_fdr,
  self_fdr_high = self_mean_fdr + 2 * self_sd_fdr)
fdr_summary <- dplyr::mutate(fdr_summary,
  FDR = factor(sprintf('%.2f', fdr_level)))
```

```{r}
ggplot(fdr_summary, aes(self_mean_fdr, mean_fdr)) +
  geom_abline(intercept = 0, slope = 1, linetype = 3) +
  geom_point(aes(color = method, shape = FDR),
    size = 9, alpha = 0.8) +
  coord_cartesian(xlim = c(-0.05, 1), ylim = c(-0.05, 1), expand = FALSE) +
  scale_color_manual(values = method_colors) +
  geom_point(aes(x, y, shape = FDR), size = 5, color = 'black',
    data = data.frame(x = c(0.01, 0.05, 0.10), y = c(0.01, 0.05, 0.10),
      FDR = factor(c('0.01', '0.05', '0.10'))), show.legend = FALSE) +
  geom_errorbar(aes(ymax = fdr_high, ymin = fdr_low, color = method),
    width = 0.01) +
  # geom_errorbar(aes(xmax = self_fdr_high, xmin = self_fdr_low, color = method),
  #   width = 0.01) +
  # theme_hp() +
  theme_cowplot(25) +
  ylab('false discovery rate') +
  xlab('self-referential false discovery rate')
  # geom_vline()
```
