---
  title: Null GEUVADIS resampling
  output:
    html_document:
      fig_height: 15
      fig_width: 10
---

# transcript level
## loading data

Take a look at `null_resampling_fdr_isoform.R` to see how this data is generated.

```{r}
suppressMessages({
library('ggplot2')
library('mamabear')
library('dplyr')
library('cowplot')
source('benchmark_methods.R')
source('gene_common.R')
})
```

```{r}
theme_hp <- function() {
  theme_cowplot(25) +
    theme(legend.key.size = unit(2, "lines"), legend.position = 'none')
}

base_dir <- '../results/final_figures'
default_extension <- '.pdf'
source('../../geuvadis/R/gene_common.R')
```


```{r}
self_benchmark <- readRDS('../results/null_resampling/isoform.rds')
```

### load Cuffdiff 2

```{r}
cuffdiff_isoform <- lapply(seq_along(self_benchmark),
  function(i) {
    path <- file.path('../results/finn_samples/null', i, 'cuffdiff')
    res <- get_cuffdiff(path)$isoform
    res <- dplyr::filter(res, status == 'OK')
  })
oracle <- self_benchmark[[1]][[1]]$oracle
cuffdiff_isoform_benchmark <- lapply(cuffdiff_isoform, function(x) {
    new_de_benchmark(list(x), 'Cuffdiff 2', oracle)
  })
```


```{r}
self_benchmark <- c(self_benchmark, list(cuffdiff_isoform_benchmark))
```

## creating the plots

### Looking at Type-I error

```{r}
self_fdr <- lapply(self_benchmark, average_sensitivity_specificity,
  use_oracle = TRUE, use_fdr = FALSE)
self_fdr <- dplyr::bind_rows(self_fdr)
self_fdr <- dplyr::mutate(self_fdr, method = sub('pval_', '', method))
```

Now let’s make some pretty labels:

```{r}
self_fdr <- dplyr::mutate(self_fdr,
  fdr_level_string = paste0('eFDR = ', sprintf('%.2f', fdr_level)))
```

```{r}
self_fdr <- dplyr::mutate(self_fdr, type1error = fp / (tp + fp + tn + fn))
```

```{r}
p <- ggplot(self_fdr, aes(method, type1error, color = method)) +
  geom_boxplot() +
  facet_wrap(~fdr_level_string, ncol = 1) +
  geom_hline(aes(yintercept = fdr_level), linetype = 2) +
  ylim(0, 1) +
  ylab('Type I error') +
  theme_hp() +
  geom_jitter() +
  scale_color_manual(values = method_colors)
p
```

```{r}
filename <- file.path(base_dir, paste0('null_resampling_isoform', default_extension))
save_plot(filename, p, base_aspect_ratio = 0.75, base_height = 15)
```

### Looking at absolute false positive

```{r}
self_fdr <- lapply(self_benchmark, average_sensitivity_specificity,
  use_oracle = TRUE, use_fdr = TRUE)
self_fdr <- dplyr::bind_rows(self_fdr)
self_fdr <- dplyr::mutate(self_fdr, method = sub('pval_', '', method))
```

Now let’s make some pretty labels:

```{r}
self_fdr <- dplyr::mutate(self_fdr,
  fdr_level_string = paste0('eFDR = ', sprintf('%.2f', fdr_level)))
```

```{r}
p <- ggplot(self_fdr, aes(method, fp, color = method)) +
  geom_boxplot(outlier.shape = NA) +
  facet_wrap(~fdr_level_string, ncol = 1) +
  # geom_hline(aes(yintercept = fdr_level), linetype = 2) +
  # ylim(0, 1) +
  # coord_trans(y = "log2") +
  scale_y_sqrt() +
  ylab('number of false positives') +
  theme_hp() +
  geom_jitter(height = 0) +
  scale_color_manual(values = method_colors)
p
```

```{r}
filename <- file.path(base_dir, paste0('null_resampling_isoform_fp', default_extension))
save_plot(filename, p, base_aspect_ratio = 0.75, base_height = 15)
```

## looking at genes

```{r}
self_benchmark <- readRDS('../results/null_resampling/gene.rds')
```

### load Cuffdiff 2

```{r}
cuffdiff_gene <- lapply(seq_along(self_benchmark),
  function(i) {
    path <- file.path('../results/finn_samples/null', i, 'cuffdiff')
    res <- get_cuffdiff(path)$gene
    res <- dplyr::filter(res, status == 'OK')
  })
oracle <- self_benchmark[[1]][[1]]$oracle
cuffdiff_gene_benchmark <- lapply(cuffdiff_gene, function(x) {
    new_de_benchmark(list(x), 'Cuffdiff 2', oracle)
  })
```

```{r}
self_benchmark <- c(self_benchmark, list(cuffdiff_gene_benchmark))
```

### Looking at absolute false positive

```{r}
self_fdr <- lapply(self_benchmark, average_sensitivity_specificity,
  use_oracle = TRUE, use_fdr = TRUE)
self_fdr <- dplyr::bind_rows(self_fdr)
self_fdr <- dplyr::mutate(self_fdr, method = sub('pval_', '', method))
```

Now let’s make some pretty labels:

```{r}
self_fdr <- dplyr::mutate(self_fdr,
  fdr_level_string = paste0('eFDR = ', sprintf('%.2f', fdr_level)))
```

```{r}
p <- ggplot(self_fdr, aes(method, fp, color = method)) +
  geom_boxplot(outlier.shape = NA) +
  facet_wrap(~fdr_level_string, ncol = 1) +
  # geom_hline(aes(yintercept = fdr_level), linetype = 2) +
  # ylim(0, 1) +
  # coord_trans(y = "log2") +
  scale_y_sqrt() +
  ylab('number of false positives') +
  theme_hp() +
  geom_jitter(height = 0) +
  scale_color_manual(values = method_colors)
p
```

```{r}
filename <- file.path(base_dir, paste0('null_resampling_gene_fp', default_extension))
save_plot(filename, p, base_aspect_ratio = 0.75, base_height = 15)
```
