---
  title: "stratifying results based on isoform complexity"
  output:
    html_document:
      fig_height: 20
      fig_width: 20
---

# preliminaries

```{r, eval=FALSE}
n_cpu <- 20
sim_name <- 'gfr_3_3_20_42_2'
sim_name <- 'gcd_3_3_20_1_2'
sim_name <- 'isoform_3_3_20_1_1'
```


```{r}
args <- commandArgs(trailingOnly = TRUE)
if (length(args) != 2) {
  stop('Usage: RScript N_CPU EXPERIMENT_STRING')
}
n_cpu <- args[1]
sim_name <- args[2]
```

The simulation string is:

```{r}
sim_name
```

We use the variable `simulation_mode` to build different axes depending on the simulation.
These axes have been tuned by hand.

```{r}
prefix <- strsplit(sim_name, '_')[[1]][1]
simulation_mode <- switch(prefix,
  'isoform' = 'independent',
  'gcd' = 'common',
  'gfr' = 'reference',
  stop(paste0('unrecognized simulation type. Please check the `sim_name`: ',
    sim_name)
  ))
```

The simulation mode is: `R simulation_mode`.

Some more global variables on where to save the figures:

```{r}
base_dir <- '../results/final_figures'
default_extension <- '.pdf'
```

A theme that seems to save well:

```{r}
theme_hp <- function() {
  theme_cowplot(25) +
    theme(legend.key.size = unit(2, "lines"))
}
```

## load packages

```{r}
suppressPackageStartupMessages({
library('cowplot')
library('data.table')
library('dplyr')
library('mamabear')
library('parallel')
})
source('gene_common.R')
source('benchmark_methods.R')
options(mc.cores = n_cpu)
```

set the number of columns for each grid

```{r}
n_grid_columns <- 2
```

ignore the following when plotting estimated fdr:

```{r}
ignore_fdr <- c('GLFC', 'LFC')
# method_colors_lfc <- c(method_colors, c(LFC = 'black', GLFC = 'dimgray'))
method_colors_lfc <- c(method_colors, c(LFC = '#4393c3', GLFC = '#fddbc7'))
```

# isoform benchmarks

first, extract all of the original data

```{r}
each_filter_benchmark <- readRDS(paste0('../results/', sim_name,
  '/isoform_benchmarks_filter_lfc_clean.rds'))
```

```{r}
original_data <- lapply(each_filter_benchmark,
  function(benchmark) {
    benchmark$original_data
  })
```

Next, let's break them into different categories and stratify them.

```{r}
transcript_gene_mapping <- get_human_gene_names()
```

If more than `maximum_isoforms`, pool them all together.

```{r}
maximum_isoforms <- 6
isoform_per_gene <- dplyr::summarize(
  dplyr::group_by(transcript_gene_mapping, ens_gene),
  n_isoform = pmin(length(target_id), maximum_isoforms))
transcript_gene_mapping <- dplyr::left_join(transcript_gene_mapping, isoform_per_gene,
  by = 'ens_gene')
```

```{r}
original_data <- lapply(original_data,
  function(benchmark) {
    lapply(benchmark, function(x) dplyr::left_join(x, transcript_gene_mapping, by = 'target_id'))
  })

unique_values <- unlist(lapply(original_data, function(bench) {
    unlist(lapply(bench, '[[', 'n_isoform'), use.names = FALSE)
  }), use.names = FALSE) %>%
  unique

unique_values <- sort(unique_values)
```

```{r}
original_oracle <- lapply(each_filter_benchmark, '[[', 'oracle')
```

```{r}
stratified_results <- lapply(unique_values,
  function(i) {
    lapply(original_data, function(benchmark) {
      lapply(benchmark, function(x) dplyr::filter(x, n_isoform == i))
    })
  })
```

```{r}
stratified_oracle <- lapply(original_oracle,
  function(oracle) {
    dplyr::left_join(oracle, transcript_gene_mapping, by = 'target_id')
  })

stratified_oracle <- lapply(unique_values,
  function(i) {
    lapply(stratified_oracle, function(oracle) {
      dplyr::filter(oracle, n_isoform == i)
    })
  })
```

```{r}
stratified_benchmark <- lapply(seq_along(stratified_results),
  function(i) {
    # i refers to the strata
    results <- stratified_results[[i]]
    lapply(seq_along(results), function(j) {
      # j refers to the jth list of results
      res <- results[[j]]
      new_de_benchmark(res, names(res), stratified_oracle[[i]][[j]],
        de_colors = method_colors_lfc, join_mode = 'union')
    })
  })
```

```{r}
stratified_fdr <- lapply(stratified_benchmark,
  function(benchmark) {
    suppressMessages(current_fdr <- get_fdr(benchmark, sim_filter = TRUE)$pvals)
  })
saveRDS(stratified_fdr, paste0('../results/', sim_name,
  '/stratified_fdr_isoform.rds'))
```

```{r eval=FALSE}
stratified_fdr <- readRDS(paste0('../results/', sim_name,
  '/stratified_fdr_isoform.rds'))
```

```{r}
current_limits <- switch(simulation_mode,
  independent = list(x = c(-0.01, 0.25), y = c(-0.01, 0.28)),
  common = list(x = c(-0.01, 0.25), y = c(-0.01, 0.20)),
  reference = list(x = c(-0.01, 0.25), y = c(-0.01, 0.25))
  )

stratified_zoom <- lapply(stratified_fdr,
  function(current_fdr) {
    tmp <- fdr_efdr_power_plot(current_fdr, start = 100, jump = 100, rank_fdr = 0.10,
      method_colors = method_colors_lfc, fdr_level_position = -0.005,
      # ignore fdr for everyone since it's computed on the whole
      ignore_estimated_fdr = unique(current_fdr$method))

    p <- tmp + theme_hp()
    p <- p + coord_cartesian(xlim = current_limits$x, ylim = current_limits$y,
      expand = FALSE)
    p
})
```

```{r}
p <- plot_grid(plotlist = stratified_zoom, labels = paste0('(',
  letters[1:length(stratified_zoom)], ')'))
p
```

```{r}
filename <- file.path(base_dir, paste0('stratify_isoform.each_filter_', sim_name,
  default_extension))
save_plot(filename, p, base_aspect_ratio = 1.6, base_height = 15)
```

```{r}
stratified_no_zoom <- lapply(stratified_fdr,
  function(current_fdr) {
    tmp <- fdr_efdr_power_plot(current_fdr, start = 500, jump = 500, rank_fdr = 0.10,
      method_colors = method_colors_lfc, fdr_level_position = -0.02,
      ignore_estimated_fdr = unique(current_fdr$method))

    p <- tmp +
      coord_cartesian(xlim = c(-0.05, 1), ylim = c(-0.05, 1), expand = FALSE) +
      theme_hp()
    p <- p + guides(colour = guide_legend(override.aes = list(size=5)))
    p <- p +
      geom_polygon(aes(x, y), alpha = 0.20,
        data = data.frame(
        x = c(0, 0, current_limits$x[2], current_limits$x[2]),
        y = c(0, current_limits$y[2], current_limits$y[2], 0)))
    p
  })
```

```{r}
p <- plot_grid(plotlist = stratified_no_zoom, labels = paste0('(',
  letters[1:length(stratified_zoom)], ')'))
p
```

```{r}
filename <- file.path(base_dir, paste0('stratify_isoform.each_filter_nozoom_', sim_name,
  default_extension))
save_plot(filename, p, base_aspect_ratio = 1.6, base_height = 15)
```

--------------------------------------------------------------------------------

# gene results

```{r}
each_filter_benchmark <- readRDS(paste0('../results/', sim_name,
  '/gene_benchmarks_filter_lfc_clean.rds'))
```

```{r}
original_data <- lapply(each_filter_benchmark,
  function(benchmark) {
    benchmark$original_data
  })
```
```{r}
gene_isoform_mapping <- dplyr::select(transcript_gene_mapping, target_id = ens_gene,
  n_isoform)
gene_isoform_mapping <- dplyr::distinct(gene_isoform_mapping)
```

```{r}
original_data <- lapply(original_data,
  function(benchmark) {
    lapply(benchmark, function(x) dplyr::left_join(x, gene_isoform_mapping, by = 'target_id'))
  })

unique_values <- unlist(lapply(original_data, function(bench) {
    unlist(lapply(bench, '[[', 'n_isoform'), use.names = FALSE)
  }), use.names = FALSE) %>%
  unique

unique_values <- sort(unique_values)
```


```{r}
original_oracle <- lapply(each_filter_benchmark, '[[', 'oracle')
```

```{r}
stratified_results <- lapply(unique_values,
  function(i) {
    lapply(original_data, function(benchmark) {
      lapply(benchmark, function(x) dplyr::filter(x, n_isoform == i))
    })
  })
```

```{r}
stratified_oracle <- lapply(original_oracle,
  function(oracle) {
    dplyr::left_join(oracle, gene_isoform_mapping, by = 'target_id')
  })

stratified_oracle <- lapply(unique_values,
  function(i) {
    lapply(stratified_oracle, function(oracle) {
      dplyr::filter(oracle, n_isoform == i)
    })
  })
```

```{r}
stratified_benchmark <- lapply(seq_along(stratified_results),
  function(i) {
    # i refers to the strata
    results <- stratified_results[[i]]
    lapply(seq_along(results), function(j) {
      # j refers to the jth list of results
      res <- results[[j]]
      new_de_benchmark(res, names(res), stratified_oracle[[i]][[j]],
        de_colors = method_colors_lfc, join_mode = 'union')
    })
  })
```


```{r}
stratified_fdr <- lapply(stratified_benchmark,
  function(benchmark) {
    suppressMessages(current_fdr <- get_fdr(benchmark)$pvals)
    current_fdr
  })
saveRDS(stratified_fdr, paste0('../results/', sim_name,
  '/stratified_fdr_gene.rds'))
```

```{r eval=FALSE}
stratified_fdr <- readRDS(paste0('../results/', sim_name,
  '/stratified_fdr_gene.rds'))
```

```{r}
current_limits <- switch(simulation_mode,
  independent = list(x = c(-0.01, 0.25), y = c(-0.01, 0.28)),
  common = list(x = c(-0.01, 0.25), y = c(-0.01, 0.20)),
  reference = list(x = c(-0.01, 0.25), y = c(-0.01, 0.25))
  )
stratified_zoom <- lapply(stratified_fdr,
  function(current_fdr) {
    tmp <- fdr_efdr_power_plot(current_fdr, start = 100, jump = 100, rank_fdr = 0.10,
      method_colors = method_colors_lfc, fdr_level_position = -0.005,
      # ignore fdr for everyone since it's computed on the whole
      ignore_estimated_fdr = unique(current_fdr$method))

    p <- tmp + theme_hp()
    p <- p + coord_cartesian(xlim = current_limits$x, ylim = current_limits$y,
      expand = FALSE)
    p <- p + guides(colour = guide_legend(override.aes = list(size=5)))
    p
})
```

```{r}
p <- plot_grid(plotlist = stratified_zoom, labels = paste0('(',
  letters[1:length(stratified_zoom)], ')'))
p
```

```{r}
filename <- file.path(base_dir, paste0('stratify_gene.each_filter_', sim_name,
  default_extension))
save_plot(filename, p, base_aspect_ratio = 1.6, base_height = 15)
```

```{r}
stratified_no_zoom <- lapply(stratified_fdr,
  function(current_fdr) {
    tmp <- fdr_efdr_power_plot(current_fdr, start = 500, jump = 500, rank_fdr = 0.10,
      method_colors = method_colors_lfc, fdr_level_position = -0.02,
      ignore_estimated_fdr = unique(current_fdr$method))

    p <- tmp +
      coord_cartesian(xlim = c(-0.05, 1), ylim = c(-0.05, 1), expand = FALSE) +
      theme_hp()
    p <- p +
      geom_polygon(aes(x, y), alpha = 0.20,
        data = data.frame(
        x = c(0, 0, current_limits$x[2], current_limits$x[2]),
        y = c(0, current_limits$y[2], current_limits$y[2], 0)))
    p <- p + guides(colour = guide_legend(override.aes = list(size=5)))
    p
  })
```

```{r}
p <- plot_grid(plotlist = stratified_no_zoom, labels = paste0('(',
  letters[1:length(stratified_no_zoom)], ')'))
p
```

```{r}
filename <- file.path(base_dir, paste0('stratify_gene.each_filter_nozoom_', sim_name,
  default_extension))
save_plot(filename, p, base_aspect_ratio = 1.6, base_height = 15)
```
